---
title: "California Education Attainment Census Data Analysis"
author: 
  - Rayan Awad Alim
  - Emily Su
  - Heyucheng Zhang
  - Maryam Ansari
  - Prankit Bhardwai
  - Luka Tosic
thanks: "Code and data are available at: https://github.com/RayanAlim/US_Census_Education_Data_Analysis/"
date: today
date-format: long
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(dplyr)
library(here)
```

# Data

We use the statistical programming language R [@citeR], IPUMS [@@ipumsusa], dpylr [@dpylr], here [@here], and tidyverse [@rohan]. The data is taken from @ipumsusa. 

## Ratio Estimator Approach 

How the ratio estimator approach works is that it estimates a population size based on a ratio of two means of information we know regarding the population. How we applied the ratio estimator is that we took the ratio of doctoral holders in California over the total number of respondents in California from our data. For each state we figured out the number of doctoral holders from our data and then divide this number with the ratio we obtained previously to get the estimated total number of respondents in each state. 

# Results

```{r}
#| include: false
#| warning: false
#| message: false

# Read the data 
data <- read_csv(here::here("data", "analysis_data", "clean_census_data.csv"))

# Filter for doctoral degree holders 
doctoral_by_state <- data %>%
  filter(EDUCD == 116) %>%
  group_by(STATEICP) %>%
  summarise(doctoral_count = n())
```

```{r}
#| echo: false
#| warning: false
#| message: false
#  total respondents in California
total_CA <- 391171

#  the number of doctoral degree holders in California
doctoral_CA <- doctoral_by_state %>%
  filter(STATEICP == 71) %>%
  pull(doctoral_count)
```

```{r}
#| echo: false
#| warning: false
#| message: false
#  ratio for cali-  (doctoral holders / total respondents)
ratio_CA <- doctoral_CA / total_CA

# est total respondents for all states
estimated_total_by_state <- doctoral_by_state %>%
  mutate(estimated_total = doctoral_count / ratio_CA)
```



# Appendix 
## How to obtain IPUMS data
In order to obtain the data from IPUMS [@@ipumsusa], first go to usa.ipums.org and then click on "Get Data" on the home page. Next click on "Select Samples", select only the 2022 ACS sample under the "USA SAMPLES" tab, and then "SUBMIT SAMPLE SELECTIONS". Then under "SELECT HARMONIZED VARIABLES" select "GEOGRAPHIC" under the "HOUSEHOLD" dropdown and choose the STATEICP variable. Under the PERSON dropdown, select EDUCATION and then the EDUC variable. After all this, click on VIEW CART on the top right and on the DATA CART Page click on "create data extract" and then on the Extract Request page make sure the data format is in the csv format. After submitting the extract on the Extract Request, create an account or log into IPUMS USA and wait for the extract to be finished and download the CSV. 


\newpage


# References


